Dirac neutrino mass generation from Majorana messenger

### 2019-08-05 draft.tex
uf

\section{Introduction}
\label{sec:intro}
%Dirac or Majorana
%TODO: check ad-hoc symmetries
The interpretation of neutrino experimental data in terms of neutrino
oscillations is compatible with both Majorana or Dirac neutrino masses
\cite{Tanabashi:2018oca}. The former possibility has received the most
attention but, given the lack of signals in neutrinoless double beta
decay experiments
\cite{KamLAND-Zen:2016pfg,Agostini:2018tnm,Aalseth:2017btx,Alduino:2017ehq,Albert:2017owj,Arnold:2016bed},
the latter cannot be dismissed.
% Dirac case
If neutrinos are Dirac particles, the Standard Model (SM) particle
content must be extended with right-handed neutrinos,
% with cosmology implications
which can increase the effective number of light neutrinos,
$N_{\text{eff}}$, until 6. Therefore, to be compatible with the
current cosmological restrictions on $N_{\text{eff}}$, the
interactions of the extra right-handed neutrinos with the primordial
plasma must be highly suppressed.

@@

%Features of the new propossed solution
In general, solutions for light Dirac neutrino masses require a
continuous symmetry to guarantee their Diracness. This symmetry is
usually identified as the $\operatorname{U}(1)_{B-L}$.  Additionally,
ad hoc discrete symmetries are invoked to forbid tree level Dirac or
Majorana mass terms for the light right-handed
neutrinos~\cite{Roncadelli:1983ty,Han:2018zcn,Wang:2017mcy}. 
%tree-level without adhoc symmetries
However, tree-level Dirac type-I seesaw with proper choices for the
$\operatorname{U}(1)_{B-L}$ charges have been shown to be consistent
without require any extra ad-hoc discrete symmetry~\cite{Ma:2014qra}.
%one-loop 
In recent works, it has been shown that even for one-loop Dirac
neutrino masses, it is possible to have $\operatorname{U}(1)_{B-L}$ as
the only extra symmetry beyond the SM~\cite{Calle:2018ovc,Bonilla:2018ynb,Saad:2019bqf}\footnote{
%extra-loops
  For extensions with only extra scalars, minimal solutions have been found with two and three loops~\cite{Saad:2019bqf}}.
%conexion with dark matter
As a bonus in this case, the new scalars and fermions circulating the
loop can be dark matter candidates with the stability of the lightest
of them guaranteed by the very same continuous symmetry.
% remark
We focus here in solutions for the radiative Dirac type-I seesaw with
Majorana mediators, which have only an extra symmetry responsible for the
Diracness of the light neutrinos, the absence of any tree-level
mass, and the existence of a dark sector constituted by the
particles circulating the loop.

@@

% We will make use of the scotogenic model~\cite{Ma:2006km}. In
% scotogenic model the neutrino masses and the DM have common
% explanation. The neutrino masses are generated by one-loop effects,
% where the candidates for DM are mediators in the loop.
%Conclusion:
In this work we propose a Dirac radiative type-I seesaw, with Majorana
fermion of the mediators in the loop. This mediators can include a DM
candidate.

@@

\section{The model}
\label{sec:Model}
%
\begin{table}[t!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline  
    Fields     & $SU(2)_L$ & $U(1)_Y $ & $U(1)_X$ \\ \hline
    % $H $  & 2  1/2  &  0 \\
    $\eta$  & 2 & 1/2  & $1$ \\
    $S$ & 1 & 0  & $2$ \\
    $\sigma$ & 1 & 0 & $3$ \\
    \hline
    % $Q_{L_{i}}$  & $(2,-1/6)$ & $0$ \\
    % $\overline{u_{R_{i}}}$ & $(1,-2/3)$ & $0$ \\
    % $\overline{d_{R_{i}}}$ & $(1,1/3)$ & $0$ \\
    % \hline
    % $L_i$  & $(2,-1/2)$ & $0$ \\
    % $\overline{e_{R_i}}$ & $(1,1)$ & $0$ \\
    $\nu_{Ri}$ & 1 & 0 & -4\\
    $\nu_{R3}$ & 1 & 0 & $5$\\
    $\psi_{R\alpha}$  & 1 & 0 & $1$ \\\hline
  \end{tabular}
  \caption{The new scalars and fermions with their respective charges. All the SM fields are neutral under the new $U(1)_X$ gauge symmetry. }
    \label{tab:partcont}
\end{table}
%
With the aim of illustrate the one-loop Dirac neutrino mass generation we consider the particle content shown in Table \ref{tab:partcont}. Specifically, we introduce three scalar fields ($\eta, \sigma$ and $S$), where only $S$ develops a nonzero vacuum expectation value (VEV), and two sets of three singlet fermions, $\nu_{Ri}$ and $\nu_{R3}$, and $\psi_{R\alpha}$, where $i=1,2$ and $\alpha=1,2,3$.   
The model also includes an additional $U(1)_X$ gauge symmetry which is used to forbid the Dirac and Majorana neutrino mass terms at tree level.
%Additionally, we add three right neutrinos, one scalar doublet, two scalar singlets and three Majorana fermion singlet under $SU(2)_L$.
The $U(1)_X$ charge for the new particles will be defined by the anomaly cancellation conditions and the gauge invariance in Yukawa and scalar interactions.
%In this model the neutrinos are Dirac fermions. Particle content and their respective charges are show in Tab.~\ref{tab:partcont}.

>>>>

\section{Introduction}
\label{sec:intro}
%Dirac or Majorana
The interpretation of neutrino experimental data in terms of neutrino
oscillations is compatible with both Majorana or Dirac neutrino masses
\cite{Tanabashi:2018oca}. The former possibility has received the most
attention but, given the lack of signals in neutrinoless double beta
decay experiments
\cite{KamLAND-Zen:2016pfg,Agostini:2018tnm,Aalseth:2017btx,Alduino:2017ehq,Albert:2017owj,Arnold:2016bed},
the latter cannot be dismissed.
% Dirac case
If neutrinos are Dirac particles, the Standard Model (SM) particle
content must be extended with right-handed neutrinos,
% with cosmology implications
which can increase the effective number of light neutrinos,
$N_{\text{eff}}$, until 6. Therefore, to be compatible with the
current cosmological restrictions on $N_{\text{eff}}$, the
interactions of the extra right-handed neutrinos with the primordial
plasma must be highly suppressed.

@@

%Features of the new propossed solution
In general, solutions for light Dirac neutrino masses require a
continuous symmetry to guarantee their Diracness. This symmetry is
usually identified as the $\operatorname{U}(1)_{B-L}$.  Additionally,
ad-hoc discrete symmetries are invoked to forbid tree level Dirac or
Majorana mass terms for the light right-handed
neutrinos~\cite{Roncadelli:1983ty,Han:2018zcn,Wang:2017mcy}. 
%tree-level without adhoc symmetries
However, tree-level Dirac type-I seesaw with proper choices for the
$\operatorname{U}(1)_{B-L}$ charges have been shown to be consistent
without require any extra ad-hoc discrete symmetry~\cite{Ma:2014qra}.
%one-loop 
In recent works, it has been shown that even for one-loop Dirac
neutrino masses, it is possible to have $\operatorname{U}(1)_{B-L}$ as
the only extra symmetry beyond the SM~\cite{Calle:2018ovc,Bonilla:2018ynb,Saad:2019bqf}\footnote{
%extra-loops
  For extensions with only extra scalars, minimal solutions have been found with two and three loops~\cite{Saad:2019bqf}}.
%conexion with dark matter
As a bonus in this case, the new scalars and fermions circulating the
loop can be dark matter candidates with the stability of the lightest
of them guaranteed by the very same continuous symmetry.
% remark
We focus here in solutions for the radiative Dirac type-I seesaw with
Majorana mediators, which have only an extra symmetry responsible for the
Diracness of the light neutrinos, the absence of any tree-level
mass, and the existence of a dark sector constituted by the
particles circulating the loop.

@@

% We will make use of the scotogenic model~\cite{Ma:2006km}. In
% scotogenic model the neutrino masses and the DM have common
% explanation. The neutrino masses are generated by one-loop effects,
% where the candidates for DM are mediators in the loop.
%Conclusion:
In this work we propose a radiative Dirac type-I seesaw, with Majorana
fermions as mediators in the loop. This full set of mediators in the
loop can include a DM candidate.

@@

\section{The model}
\label{sec:Model}
%
\begin{table}[t!]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline  
    Fields     & $SU(2)_L$ & $U(1)_Y $ & $U(1)_X$ \\ \hline
    % $H $  & 2  1/2  &  0 \\
    $\eta$  & 2 & 1/2  & $1$ \\
    $S$ & 1 & 0  & $2$ \\
    $\sigma$ & 1 & 0 & $3$ \\
    \hline
    % $Q_{L_{i}}$  & $(2,-1/6)$ & $0$ \\
    % $\overline{u_{R_{i}}}$ & $(1,-2/3)$ & $0$ \\
    % $\overline{d_{R_{i}}}$ & $(1,1/3)$ & $0$ \\
    % \hline
    % $L_i$  & $(2,-1/2)$ & $0$ \\
    % $\overline{e_{R_i}}$ & $(1,1)$ & $0$ \\
    $\nu_{Ri}$ & 1 & 0 & -4\\
    $\nu_{R3}$ & 1 & 0 & $5$\\
    $\psi_{R\alpha}$  & 1 & 0 & $1$ \\\hline
  \end{tabular}
  \caption{The new scalars and fermions with their respective charges. All the SM fields are neutral under the new $U(1)_X$ gauge symmetry. }
    \label{tab:partcont}
\end{table}
%
We extend the standard model with a spontaneously broken Abelian gauge symmetry  which guarantees the total lepton number $L$ conservation. Only the new particles including the light right handed neutrinos are charged under this symmetry. We choose the new particle set such that the following dimension six operator is genuinely realized
\begin{align}
  \mathcal{L}_{6D}=\frac{1}{\Lambda^2} \overline{L_i} H \nu_R S^2\,,
\end{align}
where $S$ is the singlet scalar field which can spontanouely break the $\operatorname{U}(1)_X$ symmetry.

With the aim to illustrate the one-loop Dirac neutrino mass generation
we consider the particle content shown in Table
\ref{tab:partcont}. Specifically, we introduce three scalar fields
($\eta, \sigma$ and $S$), where only $S$ develops a nonzero vacuum expectation
value (VEV), and two sets of three singlet fermions, $\nu_{Ri}$ and
$\nu_{R3}$, and $\psi_{R\alpha}$, where $i=1,2$ and $\alpha=1,2,3$.  The model also
includes an additional $U(1)_X$ gauge symmetry which is used to forbid
the Dirac and Majorana neutrino mass terms at tree level.
%Additionally, we add three right neutrinos, one scalar doublet, two scalar singlets and three Majorana fermion singlet under $SU(2)_L$.
The $U(1)_X$ charge for the new particles will be defined by the anomaly cancellation conditions and the gauge invariance in Yukawa and scalar interactions.
%In this model the neutrinos are Dirac fermions. Particle content and their respective charges are show in Tab.~\ref{tab:partcont}.
